The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area

### 2023-07-17 MS_ToSubmit/MS_Burdwood_foodweb.tex
MEPS format

  \journal{MEPS, Oikos, Ecological Applications, J of Applied Ecology
(?)} % Sets Journal name

@@

  \title{The complex network of trophic interactions in a subAntarctic
oceanic Marine Protected Area}
    \author[]{Tomás I. Marina\textsuperscript{a}%
  %
  }
   \ead{tomasimarina@gmail.com} 
    \author[]{Irene R. Schloss\textsuperscript{a,b,c}%
  %
  }

@@

    \author[]{Juan Martín Díaz de Astarloa\textsuperscript{e}%
  %
  }

@@

      \affiliation[1]{Centro Austral de Investigaciones Cientificas
(CADIC-CONICET), Argentina}
    \affiliation[2]{Instituto Antartico Argentino (IAA), Argentina}
    \affiliation[3]{Instituto de Ciencias Polares, Ambiente y Recursos
Naturales (ICPA), Universidad de Tierra del Fuego (UNTDF), Argentina}
    \affiliation[4]{Facultad de Ciencias Exactas y Naturales,
Universidad de Buenos Aires (UBA), Argentina}
    \affiliation[5]{Instituto de Biodiversidad y Biologia Experimental y
Aplicada (IBBEA), Universidad de Buenos Aires-CONICET, Argentina}
    \affiliation[6]{Instituto de Investigaciones Marinas y Costeras
(IIMYC), Universidad Nacional de Mar del Plata-CONICET, Argentina}
    \affiliation[7]{Wildlife Conservation Society, Argentina}
    \affiliation[8]{Instituto Nacional de Investigacion y Desarrollo
Pesquero (INIDEP), Consejo Nacional de Investigaciones Cientificas y
Tecnicas (CONICET), Argentina}
    \affiliation[9]{Facultad de Ciencias Exactas, Fisicas y Naturales,
Universidad Nacional de Cordoba (UNC), Argentina}
    \affiliation[10]{Instituto de Diversidad y Ecologia Animal
(IDEA-CONICET)}
    \affiliation[11]{Museo Argentino de Ciencias Naturales ``Bernardino
Rivadavia'', Argentina}
    \cortext[cor1]{Corresponding author}

  \begin{abstract}
  Globally, the total area of the world ocean designated under marine
  protection has increased in the recent decades. The majority of these
  Marine Protected Areas (MPAs) focus on the presence of particularly
  vulnerable, keystone, or charismatic species, large numbers of endemic
  species, and/or high biodiversity across taxonomic levels. In the
  sub-Antarctic region, the level of ocean protection is mainly
  associated to oceanic islands, except for the MPAs Namuncurá -
  Burdwood Bank I and II (MPA N-BB, \textasciitilde53º--55ºS and
  \textasciitilde56º--62ºW), which are associated to a submarine plateau
  and its southern adjacent deep slope region, respectively. Here, we
  present the first detailed analysis of the network of predator-prey
  interactions (food web) for the MPA N-BB ecosystem. We applied a
  network approach to characterise the food web in terms of complexity
  and structure, and identifying the species' role in such a framework.
  In terms of complexity, the MPA N-BB food web consisted of 1788
  predator-prey interactions and 379 species, with a link density of
  4.72 and a connectance of 0.01. In terms of structure, almost half of
  the consumers were omnivores (0.48), and the network displayed a
  small-world pattern. These findings suggest that the ecosystem might
  be vulnerable to external perturbations targeting highly connected
  species, although other properties might provide resilience and
  resistance, resulting in a rearranged structure that preserves its
  original functions. Furthermore, we identified several species as
  important in terms of different aspects of trophic structure and
  functioning, and response to perturbations. We suggest that generalist
  species, mainly fishes, play a crucial role in the ecosystem's
  bentho-pelagic coupling process and that other species, besides the
  longtail southern cod \emph{Patagonotothen ramsayi} and the Fueguian
  sprat \emph{Sprattus fuegensis}, should be considered as relevant
  energy transfers for the ecosystem. Finally, we argue that the
  diversity of species, including both the benthic and pelagic habitats,
  is responsible for seccuring the connectivity within the food web
  against perturbations, therefore contributing to the structure and
  stability of the ecosystem.
  \end{abstract}
    \begin{keyword}
    Food web \sep Complexity \sep Structure \sep Burdwood Bank \sep 
    Southwest Atlantic
  \end{keyword}

@@

The evidence of benefits provided by Marine Protected Areas (MPAs) as
well as the urgent need for ocean protection have driven an
unprecedented increase in the number of MPAs worldwide in recent years
(Roberts et al., 2017; Sala et al., 2018). Globally, the total area of
the world ocean designated under marine protection adds up to
approximately 29,600,000 km\textsuperscript{2}, distributed across
nearly 18,444 MPAs and covering 8.16\% of the ocean's surface (IUCN \&
UNEP-WCMC, 2023), and therefore approaching the 10\% goal of the
Convention of Biological Diversity (Secretariat of the convention on
biological diversity, 2004). Despite this progress, recent reports have
shown that actual protection has been overestimated because it includes
areas that are not yet effectively protected (only declared) as well as
areas that allow significant extractive activities (Sala et al., 2018).

In the sub-Antarctic region, the level of ocean protection is mainly
associated to oceanic islands, such as the South Georgias and South
Sandwich, Bouvet, Prince Edward, and Macquarie islands (IUCN \&
UNEP-WCMC, 2023). Interestingly, the case of the MPAs Namuncurá -
Burdwood Bank I and II (MPA N-BB, \textasciitilde53º--55ºS
\textasciitilde56º--62ºW), which is the focus of this work, is unique
since these MPAs are associated to a submarine plateau and its southern
adjacent deep slope region, respectively (Falabella, 2017; Schejter et
al., 2020). In addition, such MPAs are part of a network of protected
areas in the sub-Antarctic area (jointly with MPA Yaganes) that aims to
protect this southern region in order to contribute to global ocean
health.

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, large numbers (or proportions) of
endemic species, and/or high biodiversity across taxonomic levels (Hogg
et al., 2016). Indeed, the MPA N-BB was created to protect a potentially
sensitive and biodiverse benthic habitat that was only barely known
(Falabella, 2017; Schejter et al., 2016). The benthic community is
featured by high biomass of vulnerable and fragile species (mainly
Porifera, Bryozoa and Cnidaria) that considered with their environment
meet the characteristics of vulnerable marine ecosystems (Schejter \&
Albano, 2021), here defined as sites that present densities of Indicator
Taxa of \textgreater{} 10 kg per 1200 m\textsuperscript{2} (Commission
for the Conservation of Antarctic Marine Living Resources (CCAMLR),
2009). Also, the benthic realm provides habitat to several small-sized
species (López-Gappa et al., 2018; Martin Sirito, 2019; Schejter \&
Bremec, 2019), and has an important role in the life history of fishes
as a food source, refuge and nursery area (Covatti Ale et al., 2022;
Delpiani et al., 2020; Fischer et al., 2022; Florencia et al., 2023;
García Alonso et al., 2018; Matusevich, 2022; Troccoli et al., 2020;
Vazquez et al., 2018). The maintenance of this singular community is
related to local and regional oceanographic processes, including the
circulation of the rich Malvinas (Falkland) current in the area
(Guerrero et al., 1999; Piola \& Gordon, 1989) and the upwelling and
mixing phenomena (Matano et al., 2019). The input of nutrients from the
Malvinas (Falkland) current also supports a diverse plankton community
(Guinder et al., 2020).

@@

High-latitude marine ecosystems, such as the MPA N-BB, are complex
systems in terms of biodiversity and ecological interactions (Cordone et
al., 2020; Day et al., 2013; Kortsch et al., 2019; Trathan et al.,
2021). Although there is a robust knowledge about the complexity
considering the richness of the benthic and plankton communities in the
MPA N-BB ecosystem (Administración de Parques Nacionales, 2022; Guinder
et al., 2020; Schejter et al., 2016, 2020), a better understanding of
species interactions' complexity and structure is needed. This aspect
can be tackled by analysing one of the most frequent relationships
between species: the predator-prey interaction (Bascompte, 2009). The
sum of predator-prey or trophic interactions of a particular region is
referred to as a food web, representing the roadmap for matter and
energy flow in an ecosystem. In recent years, network approaches have
been successfully applied to study complex high-latitude marine
ecosystems, improving our knowledge on structure, functioning, and
response to environmental/anthropogenic changes (Cordone et al., 2018;
Funes et al., 2022; Kortsch et al., 2015; Marina et al., 2023). Among
anthropogenic threats, it is worth mentioning that contaminants like
mercury and microplastics have been recently reported as important
threats to the MPA N-BB region (Cossi et al., 2021; Di Mauro et al.,
2022; Fioramonti et al., 2022); also fishing vessels are allowed to
operate in the western section of the MPA N-BB (i.e.~Marine National
Reserve category), altering the stocks of commercially important fish
species (Administración de Parques Nacionales, 2022; Martínez et al.,
2021). Moreover, there is a potential hazard related to the effects of
offshore activities (exploration and explotation) to the west of the MPA
N-BB (Administración de Parques Nacionales, 2022).

@@

The MPAs Namuncurá - Burdwood Bank I and II, created by National Laws
26.875 in 2013 and 27.490 in 2017, comprise a shallow submarine plateau
called Burdwood Bank (BB) and a deep slope that reaches 4000 m in depth,
N-BB I and N-BB II, respectively (Administración de Parques Nacionales,
2022; Tombesi et al., 2020) (Figure 1). They are located 150 km east of
Isla de los Estados and 200 km south of Malvinas/Falkland Islands. The
MPA N-BB I comprises nearly 28,900 km\textsuperscript{2} circumscribed
by the 200 m isobath, between \textasciitilde54º--55ºS and
\textasciitilde56º--62ºW, with a slight slope extended nearly 370 km
east--west. Physical features in the BB are fairly stable, with salinity
averaging 34 all year round and temperature ranging between 4 and 8ºC
(Acha et al., 2004; Guerrero et al., 1999; Piola \& Falabella, 2009).
The BB is surrounded by steep flanks of up to 4000 m depth through which
strong currents circulate (Matano et al., 2019; Piola \& Gordon, 1989;
Reta, 2014). The N-BB II includes such a deep slope, protecting about
32,000 km\textsuperscript{2} (\textasciitilde55º-56ºS,
\textasciitilde58º-62ºW). Intense upwelling and mixing occur in relation
with the slope, entraining deep nutrient-rich waters into the photic
layer (Matano et al., 2019; Piola \& Falabella, 2009) and resulting in a
fairly homogeneous water column both spatially and temporally (Glorioso
\& Flather, 1995; Guerrero et al., 1999; Matano et al., 2019).

@@

The network-level analysis aims to characterise the food web in terms of
complexity and structure. For this, we calculated several network
properties commonly used to describe empirical food webs (Pascual \&
Dunne, 2005): (1) number of species S; (2) number of interactions or
links L; (3) link density L/S; (4) connectance L/S\^{}2; (5) omnivory
Omn; and (6) small-world pattern. In order to explore the small-world
phenomenon, we analysed the characteristic path length (CPL) and the
clustering coefficient (CC). The CPL is the average shortest path length
between all pairs of nodes (Watts \& Strogatz, 1998). Here, CPL was
calculated as the average number of nodes in the shortest path
\(CPL_{Min} (i,j)\) between all pairs of nodes \(S(i,j)\) in a network
averaged over \(S(S-1)/2\) nodes:

@@

\[
CC_i = \frac{2E_i}{K_i(K_i-1)}
\] where \(E_i\) is the effective number of interactions between \(K_i\)
nearest-neighbour nodes of node \(i\) and the maximal possible number of
such interactions (Newman, 2003). To test whether the food web presented
the small-world pattern, we compared the empirical values of CPL and CC
with those resulting from 1000 randomly generated networks with the same
size (S) and number of interactions (L), following the method proposed
by Marina, Saravia, et al. (2018).

@@

The species-level analysis aims to describe the species' role in the
food web. For this, we considered the following properties: betweenness
Btw, closeness Cl, trophic similarity TS, topological role TR, and
trophic level TL (Table 1). Topological roles refer to the fact that
food webs tend to naturally organize in non-random, modular patterns,
where modules are defined as a group of species that interact more
frequently among themselves than with species that are not members of
the module (Guimerà \& Nunes Amaral, 2005). Species can play different
roles in this respect, according to the pattern of interactions within
their own module and/or across modules. We computed the topological role
for each species, classified as module hub, species with a relatively
high number of interactions, but most within its own module; module
specialist, species with relatively few interactions and most within its
own module; module connector, species with relatively few interactions
mainly between modules; and network connector, species with high
connectivity between and within modules (Guimerà \& Nunes Amaral, 2005).

@@

All network analyses and graphs were performed in R version 4.2.2 (Team,
2022), mainly using `igraph' (Csardi \& Nepusz, 2006) and `multiweb'
(Saravia, 2022) packages. The source code and data are available at
https://github.com/TomasMarina/Banco-Burdwood.

@@

Food web connectance is a feature that resumes the complexity of the
network, but more importantly, it is an emergent property of pairwise
species interactions (Poisot \& Gravel, 2014). It contains information
regarding how interactions within an ecological network are distributed
and predicts reasonably well key dynamical properties of ecological
networks (Dunne et al., 2002c). Complex marine food webs (i.e.~with more
than 25 trophic species) show connectance values ranging from 0.01 -
0.27 (Marina, Saravia, et al., 2018). In particular, food webs from
high-latitude regions tend to exhibit a connectance closer to the
minimum (between 0.01 and 0.05) (Kortsch et al., 2015; Rodriguez et al.,
2022; Santana et al., 2013). Whether food webs display a low or a high
connectance helps to better comprehend ecosystem's synthetic properties
like robustness. In this sense, empirical analyses support the notion
that highly-connected ecological networks are robust against external
perturbations such as the introduction of new (e.g., invasive) species
(Smith-Ramesh et al., 2017) as well as species removal (e.g., local
extinction) (Dunne et al., 2002a; Montoya \& Solé, 2003). The
connectance of the food web of the MPA Namuncurá - Burdwood Bank (0.01)
is one of the lowest reported so far for these regions; in particular,
it appears to be much lower than that of Beagle Channel (0.05), an
adjacent coastal area (Rodriguez et al., 2022).

The degree distribution, the distribution of the number of interactions
per species, is the core of the structure of species interactions, which
influences the opportunities for multiple species to persist in the long
term and, therefore, their coexistence (Godoy et al., 2018). The food
web for the MPA N-BB presents an asymmetric degree distribution. This
pattern was identified at different levels of analysis: food web,
predator, prey, and functional group. Such asymmetry is a well-known
feature in empirical complex food webs in particular (Dunne et al.,
2002c; Montoya \& Solé, 2003; Stouffer et al., 2005), and has received
great attention in complex networks in general (Albert \& Barabási,
2002; Newman, 2003). The degree distribution affects the resilience of
complex food webs against random failures and pressure on a particular
component of the web: food webs showing right-skewed distributions, like
the one described in this study, are more vulnerable to the removal of
the most connected species or hubs, with the potential of producing
secondary extinctions and a catastrophic fragmentation of the network
(Albert et al., 2000; Dunne et al., 2002a; Eklöf \& Ebenman, 2006).

It is suggested that the small-world pattern, i.e., a network with short
path length and high clustering coefficient, is not frequent in complex
marine food webs, mainly due to a low clustering coefficient compared to
random networks (Dunne et al., 2002b; Marina, Saravia, et al., 2018).
However, the food web of the MPA N-BB does display a small-world
pattern. Consequences of this could be of great importance in
recognizing species evolutionary paths and the vulnerability to
perturbations (Montoya \& Solé, 2002). On the one hand, a short path
length implies a rapid spread of an impact (e.g., contaminant,
population fluctuation, local extinction) throughout the network but, at
the same time, more potentially adaptive dynamics in the face of
external perturbations (Montoya \& Solé, 2002; Williams et al., 2002).
On the other hand, a high clustering coefficient indicates the formation
of subnetworks composed only by the neighbours of particular species.
This translates into a greater resistance of the network due to the
confinement of perturbations mainly within subnetworks and not spreading
between them (Heer et al., 2020; Kortsch et al., 2019). Overall, a
small-world topology provides ecological networks with greater
resilience and resistance (Bornatowski et al., 2017; Dormann et al.,
2017).

Omnivory acts as a buffer to changes as the ecosystem presents
alternative energy pathways in the face of perturbations, i.e., reducing
the risk of cascading extinctions following the primary loss of species
(Borrvall et al., 2000). Omnivores are species able to adapt faster and
to a broader range of environmental conditions by changing their
foraging habits to feed on the most abundant prey (Fagan, 1997).
Furthermore, omnivory can be analysed from the interaction point of
view: theoretical studies have identified omnivorous interactions as a
possible candidate for a keystone interaction, sensu Kadoya et al.
(2018), highlighting the importance of omnivory in stabilizing food web
dynamics (McCann \& Hastings, 1997; Neutel et al., 2002). The high
proportion of omnivory in the food web of the MPA N-BB suggests that the
network might be robust to variations in prey abundances, which could
increase food web's persistence and stability (Stouffer \& Bascompte,
2010).

In summary, the food web of the MPA N-BB presents a combination of
network properties that makes it unique in terms of network resolution,
complexity, and structural pattern. All this suggests that the food web
might be fragile to external perturbations targeting highly connected
species, which in turn coincides to be commercial exploited species as
fishes (Laptikhovsky et al., 2013; Martínez et al., 2015; Winter \&
Arkhipkin, 2023). However, structural properties might provide
resilience and resistance with the final outcome of a rearranged
structure maintaining its functions.

@@

The degree distribution allows identifying important species, such as
potential keystone species (i.e.~highly connected) (Dunne et al., 2002a;
Solé \& Montoya, 2001), generalist/specialist species, and dominant food
sources (Kondoh et al., 2010).

>>>>

@@

  \title{The complex network of trophic interactions in a subAntarctic
oceanic Marine Protected Area}
    \author[]{Tomás I. Marina\textsuperscript{a}%
  \corref{cor1}%
  }
   \ead{tomasimarina@gmail.com} 
    \author[]{Irene R. Schloss\textsuperscript{a,b,c}%
  %
  }

@@

    \author[]{Juan Martín Díaz de Astarloa\textsuperscript{f}%
  %
  }

@@

      \affiliation[1]{Centro Austral de Investigaciones Científicas
(CADIC-CONICET), Ushuaia, Argentina}
    \affiliation[2]{Instituto Antártico Argentino (IAA), CABA,
Argentina}
    \affiliation[3]{Instituto de Ciencias Polares, Ambiente y Recursos
Naturales (ICPA), Universidad de Tierra del Fuego (UNTDF), Ushuaia,
Argentina}
    \affiliation[4]{Facultad de Ciencias Exactas y Naturales,
Universidad de Buenos Aires (UBA), CABA, Argentina}
    \affiliation[5]{Instituto de Biodiversidad y Biología Experimental y
Aplicada (IBBEA), Universidad de Buenos Aires-CONICET, CABA, Argentina}
    \affiliation[6]{Instituto de Investigaciones Marinas y Costeras
(IIMYC), Universidad Nacional de Mar del Plata-CONICET, Mar del Plata,
Argentina}
    \affiliation[7]{Wildlife Conservation Society, Argentina}
    \affiliation[8]{Instituto Nacional de Investigación y Desarrollo
Pesquero (INIDEP), Consejo Nacional de Investigaciones Científicas y
Técnicas (CONICET), Mar del Plata, Argentina}
    \affiliation[9]{Facultad de Ciencias Exactas, Físicas y Naturales,
Universidad Nacional de Córdoba (UNC), Córdoba, Argentina}
    \affiliation[10]{Instituto de Diversidad y Ecología Animal
(IDEA-CONICET), Córdoba, Argentina}
    \affiliation[11]{Museo Argentino de Ciencias Naturales ``Bernardino
Rivadavia'', CABA, Argentina}
    \cortext[cor1]{Corresponding author}

  \begin{abstract}
  Abstract: The total area of the world ocean designated under marine
  protection has increased recently. Most Marine Protected Areas (MPAs)
  target vulnerable, keystone, charismatic and or endemic species. In
  the sub-Antarctic, ocean protection is associated to oceanic
  islands,except for MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB,
  \textasciitilde53º--55ºS and \textasciitilde56º--62ºW), which are
  associated to a submarine plateau and its southern adjacent deep
  slope, respectively. Here, we present the first analysis of the
  network of predator-prey interactions for the MPA N-BB. We applied a
  network approach to characterise the complexity and structure of the
  food web, and identify the species' role in such a framework. The MPA
  N-BB food web consisted of 1788 interactions and 379 species, with a
  connectance of 0.01. Almost half of the consumers were omnivores
  (0.48), and the network displayed a small-world pattern. These suggest
  that the ecosystem might be vulnerable to perturbations targeting
  highly connected species, although other properties might provide
  resilience and resistance, resulting in a rearranged structure that
  preserves its original functions. Several species arose as important
  in terms of different aspects of trophic structure and functioning,
  and response to perturbations. Generalist species, mainly fishes, play
  a crucial role in the ecosystem's benthopelagic coupling and should be
  considered as relevant energy transfers for the ecosystem. We argue
  that the diversity of species, including both the benthic and pelagic
  habitats, is responsible for securing the connectivity within the food
  web against perturbations, therefore contributing to the structure and
  stability of the ecosystem.
  \end{abstract}
    \begin{keyword}
    Food Web \sep Complexity \sep Structure \sep Burdwood Bank \sep 
    Southwest Atlantic
  \end{keyword}

@@

The evidence of benefits provided by Marine Protected Areas (MPAs) as
well as the urgent need for ocean protection have driven an
unprecedented increase in the number of MPAs worldwide in recent years
(Roberts et al., 2017; Sala et al., 2018). Globally, the total area of
the world ocean designated under marine protection adds up to
approximately 29,600,000 km\textsuperscript{2}, distributed across
nearly 18,444 MPAs and covering 8.16\% of the ocean's surface (IUCN and
UNEP-WCMC, 2023), and therefore approaching the 10\% goal of the
Convention of Biological Diversity (Secretariat of the convention on
biological diversity, 2004). Despite this progress, recent reports have
shown that actual protection has been overestimated because it includes
areas that are not yet effectively protected (only declared) as well as
areas that allow significant extractive activities (Sala et al., 2018).

In the sub-Antarctic region, the level of ocean protection is mainly
associated to oceanic islands, such as the South Georgias and South
Sandwich, Bouvet, Prince Edward, and Macquarie islands (IUCN and
UNEP-WCMC, 2023). Interestingly, the case of the MPAs Namuncurá -
Burdwood Bank I and II (MPA N-BB, \textasciitilde53º--55ºS
\textasciitilde56º--62ºW), which is the focus of this work, is unique
since these MPAs are associated to a submarine plateau and its southern
adjacent deep slope region, respectively (Falabella, 2017; Schejter et
al., 2020). In addition, such MPAs are part of a network of protected
areas in the sub-Antarctic area (jointly with MPA Yaganes) that aims to
protect this southern region in order to contribute to global ocean
health.

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, large numbers (or proportions) of
endemic species, and/or high biodiversity across taxonomic levels (Hogg
et al., 2016). Indeed, the MPA N-BB was created to protect a potentially
sensitive and biodiverse benthic habitat that was only barely known
(Falabella, 2017; Schejter et al., 2016). The benthic community is
featured by high biomass of vulnerable and fragile species (mainly
Porifera, Bryozoa and Cnidaria) that considered with their environment
meet the characteristics of vulnerable marine ecosystems (Schejter and
Albano, 2021), here defined as sites that present densities of Indicator
Taxa of \textgreater{} 10 kg per 1200 m\textsuperscript{2} (Commission
for the Conservation of Antarctic Marine Living Resources (CCAMLR),
2009). Also, the benthic realm provides habitat to several small-sized
species (López-Gappa et al., 2018; Martin Sirito, 2019; Schejter and
Bremec, 2019), and has an important role in the life history of fishes
as a food source, refuge and nursery area (Covatti Ale et al., 2022;
Delpiani et al., 2020; Fischer et al., 2022; Florencia et al., 2023;
García Alonso et al., 2018; Matusevich, 2022; Troccoli et al., 2020;
Vazquez et al., 2018). The maintenance of this singular community is
related to local and regional oceanographic processes, including the
circulation of the rich Malvinas (Falkland) current in the area
(Guerrero et al., 1999; Piola and Gordon, 1989) and the upwelling and
mixing phenomena (Matano et al., 2019). The input of nutrients from the
Malvinas (Falkland) current also supports a diverse plankton community
(Guinder et al., 2020).

@@

High-latitude marine ecosystems, such as the MPA N-BB, are complex
systems in terms of biodiversity and ecological interactions (Cordone et
al., 2020; Day et al., 2013; Kortsch et al., 2019; Trathan et al.,
2021). Although there is a robust knowledge about the complexity
considering the richness of the benthic and plankton communities in the
MPA N-BB ecosystem (Administración de Parques Nacionales, 2022; Guinder
et al., 2020; Schejter et al., 2020, 2016), a better understanding of
species interactions' complexity and structure is needed. This aspect
can be tackled by analysing one of the most frequent relationships
between species: the predator-prey interaction (Bascompte, 2009). The
sum of predator-prey or trophic interactions of a particular region is
referred to as a food web, representing the roadmap for matter and
energy flow in an ecosystem. In recent years, network approaches have
been successfully applied to study complex high-latitude marine
ecosystems, improving our knowledge on structure, functioning, and
response to environmental/anthropogenic changes (Cordone et al., 2018;
Funes et al., 2022; Kortsch et al., 2015; Marina et al., 2023). Among
anthropogenic threats, it is worth mentioning that contaminants like
mercury and microplastics have been recently reported as important
threats to the MPA N-BB region (Cossi et al., 2021; Di Mauro et al.,
2022; Fioramonti et al., 2022); also fishing vessels are allowed to
operate in the western section of the MPA N-BB (i.e.~Marine National
Reserve category), altering the stocks of commercially important fish
species (Administración de Parques Nacionales, 2022; Martínez et al.,
2021). Moreover, there is a potential hazard related to the effects of
offshore activities (exploration and explotation) to the west of the MPA
N-BB (Administración de Parques Nacionales, 2022).

@@

The MPAs Namuncurá - Burdwood Bank I and II, created by National Laws
26.875 in 2013 and 27.490 in 2017, comprise a shallow submarine plateau
called Burdwood Bank (BB) and a deep slope that reaches 4000 m in depth,
N-BB I and N-BB II, respectively (Administración de Parques Nacionales,
2022; Tombesi et al., 2020) (Figure 1). They are located 150 km east of
Isla de los Estados and 200 km south of Malvinas/Falkland Islands. The
MPA N-BB I comprises nearly 28,900 km\textsuperscript{2} circumscribed
by the 200 m isobath, between \textasciitilde54º--55ºS and
\textasciitilde56º--62ºW, with a slight slope extended nearly 370 km
east--west. Physical features in the BB are fairly stable, with salinity
averaging 34 all year round and temperature ranging between 4 and 8ºC
(Acha et al., 2004; Guerrero et al., 1999; Piola and Falabella, 2009).
The BB is surrounded by steep flanks of up to 4000 m depth through which
strong currents circulate (Matano et al., 2019; Piola and Gordon, 1989;
Reta, 2014). The N-BB II includes such a deep slope, protecting about
32,000 km\textsuperscript{2} (\textasciitilde55º-56ºS,
\textasciitilde58º-62ºW). Intense upwelling and mixing occur in relation
with the slope, entraining deep nutrient-rich waters into the photic
layer (Matano et al., 2019; Piola and Falabella, 2009) and resulting in
a fairly homogeneous water column both spatially and temporally
(Glorioso and Flather, 1995; Guerrero et al., 1999; Matano et al.,
2019).

@@

The network-level analysis aims to characterise the food web in terms of
complexity and structure. For this, we calculated several network
properties commonly used to describe empirical food webs (Pascual and
Dunne, 2005): (1) number of species S; (2) number of interactions or
links L; (3) link density L/S; (4) connectance L/S\^{}2; (5) omnivory
Omn; and (6) small-world pattern. In order to explore the small-world
phenomenon, we analysed the characteristic path length (CPL) and the
clustering coefficient (CC). The CPL is the average shortest path length
between all pairs of nodes (Watts and Strogatz, 1998). Here, CPL was
calculated as the average number of nodes in the shortest path
\(CPL_{Min} (i,j)\) between all pairs of nodes \(S(i,j)\) in a network
averaged over \(S(S-1)/2\) nodes:

@@

\[
CC_i = \frac{2E_i}{K_i(K_i-1)}
\] where \(E_i\) is the effective number of interactions between \(K_i\)
nearest-neighbour nodes of node \(i\) and the maximal possible number of
such interactions (Newman, 2003). To test whether the food web presented
the small-world pattern, we compared the empirical values of CPL and CC
with those resulting from 1000 randomly generated networks with the same
size (S) and number of interactions (L), following the method proposed
by Marina et al. (2018b).

@@

The species-level analysis aims to describe the species' role in the
food web. For this, we considered the following properties: betweenness
Btw, closeness Cl, trophic similarity TS, topological role TR, and
trophic level TL (Table 1). Topological roles refer to the fact that
food webs tend to naturally organize in non-random, modular patterns,
where modules are defined as a group of species that interact more
frequently among themselves than with species that are not members of
the module (Guimerà and Nunes Amaral, 2005). Species can play different
roles in this respect, according to the pattern of interactions within
their own module and/or across modules. We computed the topological role
for each species, classified as module hub, species with a relatively
high number of interactions, but most within its own module; module
specialist, species with relatively few interactions and most within its
own module; module connector, species with relatively few interactions
mainly between modules; and network connector, species with high
connectivity between and within modules (Guimerà and Nunes Amaral,
2005).

@@

All network analyses and graphs were performed in R version 4.2.2 (Team,
2022), mainly using `igraph' (Csardi and Nepusz, 2006) and `multiweb'
(Saravia, 2022) packages. The source code and data are available at
https://github.com/TomasMarina/Banco-Burdwood.

@@

Food web connectance is a feature that resumes the complexity of the
network, but more importantly, it is an emergent property of pairwise
species interactions (Poisot and Gravel, 2014). It contains information
regarding how interactions within an ecological network are distributed
and predicts reasonably well key dynamical properties of ecological
networks (Jennifer A. Dunne et al., 2002a). Complex marine food webs
(i.e.~with more than 25 trophic species) show connectance values ranging
from 0.01 - 0.27 (Marina et al., 2018b). In particular, food webs from
high-latitude regions tend to exhibit a connectance closer to the
minimum (between 0.01 and 0.05) (Kortsch et al., 2015; Rodriguez et al.,
2022; Santana et al., 2013). Whether food webs display a low or a high
connectance helps to better comprehend ecosystem's synthetic properties
like robustness. In this sense, empirical analyses support the notion
that highly-connected ecological networks are robust against external
perturbations such as the introduction of new (e.g., invasive) species
(Smith-Ramesh et al., 2017) as well as species removal (e.g., local
extinction) (Jennifer A. Dunne et al., 2002b; Montoya and Solé, 2003).
The connectance of the food web of the MPA Namuncurá - Burdwood Bank
(0.01) is one of the lowest reported so far for these regions; in
particular, it appears to be much lower than that of Beagle Channel
(0.05), an adjacent coastal area (Rodriguez et al., 2022).

The degree distribution, the distribution of the number of interactions
per species, is the core of the structure of species interactions, which
influences the opportunities for multiple species to persist in the long
term and, therefore, their coexistence (Godoy et al., 2018). The food
web for the MPA N-BB presents an asymmetric degree distribution. This
pattern was identified at different levels of analysis: food web,
predator, prey, and functional group. Such asymmetry is a well-known
feature in empirical complex food webs in particular (Jennifer A. Dunne
et al., 2002a; Montoya and Solé, 2003; Stouffer et al., 2005), and has
received great attention in complex networks in general (Albert and
Barabási, 2002; Newman, 2003). The degree distribution affects the
resilience of complex food webs against random failures and pressure on
a particular component of the web: food webs showing right-skewed
distributions, like the one described in this study, are more vulnerable
to the removal of the most connected species or hubs, with the potential
of producing secondary extinctions and a catastrophic fragmentation of
the network (Albert et al., 2000; Jennifer A. Dunne et al., 2002b; Eklöf
and Ebenman, 2006).

It is suggested that the small-world pattern, i.e., a network with short
path length and high clustering coefficient, is not frequent in complex
marine food webs, mainly due to a low clustering coefficient compared to
random networks (Jennifer A. Dunne et al., 2002c; Marina et al., 2018b).
However, the food web of the MPA N-BB does display a small-world
pattern. Consequences of this could be of great importance in
recognizing species evolutionary paths and the vulnerability to
perturbations (Montoya and Solé, 2002). On the one hand, a short path
length implies a rapid spread of an impact (e.g., contaminant,
population fluctuation, local extinction) throughout the network but, at
the same time, more potentially adaptive dynamics in the face of
external perturbations (Montoya and Solé, 2002; Williams et al., 2002).
On the other hand, a high clustering coefficient indicates the formation
of subnetworks composed only by the neighbours of particular species.
This translates into a greater resistance of the network due to the
confinement of perturbations mainly within subnetworks and not spreading
between them (Heer et al., 2020; Kortsch et al., 2019). Overall, a
small-world topology provides ecological networks with greater
resilience and resistance (Bornatowski et al., 2017; Dormann et al.,
2017).

Omnivory acts as a buffer to changes as the ecosystem presents
alternative energy pathways in the face of perturbations, i.e., reducing
the risk of cascading extinctions following the primary loss of species
(Borrvall et al., 2000). Omnivores are species able to adapt faster and
to a broader range of environmental conditions by changing their
foraging habits to feed on the most abundant prey (Fagan, 1997).
Furthermore, omnivory can be analysed from the interaction point of
view: theoretical studies have identified omnivorous interactions as a
possible candidate for a keystone interaction, sensu Kadoya et al.
(2018), highlighting the importance of omnivory in stabilizing food web
dynamics (McCann and Hastings, 1997; Neutel et al., 2002). The high
proportion of omnivory in the food web of the MPA N-BB suggests that the
network might be robust to variations in prey abundances, which could
increase food web's persistence and stability (Stouffer and Bascompte,
2010).

In summary, the food web of the MPA N-BB presents a combination of
network properties that makes it unique in terms of network resolution,
complexity, and structural pattern. All this suggests that the food web
might be fragile to external perturbations targeting highly connected
species, which in turn coincides to be commercial exploited species as
fishes (Laptikhovsky et al., 2013; Martínez et al., 2015; Winter and
Arkhipkin, 2023). However, structural properties might provide
resilience and resistance with the final outcome of a rearranged
structure maintaining its functions.

@@

The degree distribution allows identifying important species, such as
potential keystone species (i.e.~highly connected) (Jennifer A. Dunne et
al., 2002b; Solé and Montoya, 2001), generalist/specialist species, and
dominant food sources (Kondoh et al., 2010).

We have identified that most of the consumers in the food web of the MPA
N-BB either have a narrow diet or are specialists, while few present a
broad or generalist diet. The most evident generalist species are
\emph{Patagonotothen guntheri} (Covatti Ale et al., 2022), \emph{P.
ramsayi} (Fischer et al., 2022), juveniles of \emph{Dissostichus
eleginoides} (Troccoli et al., 2020), \emph{Bathyraja brachyurops}
(Belleggia et al., 2008), and \emph{B. griseocauda} (Bellegia et al.,
2014), with more than 25 potential prey. Since these species present
mid-trophic positions in the food web (with the exception of adults of
\emph{Dissostichus eleginoides} that are top predators), acting as
predator and prey, they might be important links between lower and
higher trophic levels. This result is in agreement with the sole
analysis, using stable isotopes, that exists so far for the trophic
structure of the MPA N-BB (Riccialdelli et al., 2020), and resembles
other high-latitude marine systems of the Southwest Atlantic and
Antarctic regions (Arkhipkin and Laptikhovsky, 2013; Marina et al.,
2018a). The importance of these particular generalist species also
arises since they feed in the benthic and pelagic habitats (Covatti Ale
et al., 2022; Fischer et al., 2022; Troccoli et al., 2020), linking
these realms and contributing to the vertical carbon flow.
